The state has broad powers to regulate land use. What if a state regulator
agrees to limit regulation in return for a strip of land? The transaction is
voluntary, but had the state just taken the land, it would have had to pay just
compensation. Since the government isn't obligated to allow the project, doesn't
the offer leave the landowner better off? Or is this a form of extortion? 

These types of conditional grants of permits or other dispensations under land
use regulations are called \textbf{exactions}. In \textit{Nollan v. California
Coastal Commission}, 483 U.S. 825 (1987), the Court opened the door to closer
scrutiny of these exchanges, declaring that permit conditions must serve the
same purpose as the reason to withhold permission in the first place. Absent an
``essential nexus'' between the condition and the reason for the restriction,
the demand is a taking. \textit{Id.} at 837. 

\textit{Nollan} involved a permit request to tear down and rebuild a beachfront
house. Because the project would reduce views of the ocean, the California
Coastal Commission conditioned the permit on the Nollans' granting a public
easement on their property to access the beach. The Court ruled this condition
lacked the requisite nexus. To the extent that the project would impair
sightlines to the beach, the state could condition permit approval on
ameliorative steps, like size restrictions, limits on fencing, or provision of a
platform to improve the public's view of the beach. But the majority found it
``quite impossible to understand how a requirement that people already on the
public beaches be able to walk across the Nollans' property reduces any
obstacles to viewing the beach created by the new house.'' \textit{Id.} at 838.

The outcome of \textit{Nollan} rested on the majority's conclusion that there
was \textit{no} logical relationship between the condition demanded by the
Coastal Commission and the harm it claimed to be regulating: a right to cross
the Nollan's land wouldn't improve the public's view of the beach from behind
their house. What if there is \textit{some} logical relationship, but it is (at
least arguably) somewhat attenuated?

