\expected{murr-v-wisconsin}

\item How confident are you in your ability to apply the majority's test? Is it
a fair objection that the analysis inappropriately smuggles considerations
applicable to substantive takings analysis into the distinct preliminary issue
of defining the private property in question? Or are preexisting lot lines too
arbitrary? 

\item If the dissent had prevailed, would property owners have an incentive to
break their parcels up into smaller lots to increase the likelihood of
prevailing in a takings case? Chief Justice Roberts asserts that such efforts
would be easy to ``detect and disarm.'' But what would be the mechanism for
doing so? And what if a piece of property had been owned and transferred as an
undifferentiated whole for decades, but nevertheless encompassed portions that
had once been sold as individual, smaller lots? Should they be treated
separately in a regulatory takings case?


\defwebsite{hills-half-hearted}{
author=Rick Hills,
title=A Half-Hearted Two Cheers for the Victory of Federalism over Property
Rights in Murr v. Wisconsin,
date=jun 23 2017,
journal=PrawfsBlawg,
url=http://prawfsblawg.blogs.com/prawfsblawg/2017/06/a-half-hearted-two-cheers-for-the-victory-of-federalism-over-property-rights-in-murr-v-wisconsin.html,
}

\defwebsite{somin-more-murr}{
author=Ilya Somin,
title=More on Murr---a Response to Rick Hills,
date=Jun 24 2017,
journal=The Washington Post,
url=https://www.washingtonpost.com/news/volokh-conspiracy/wp/2017/06/23/more-on-murr-a-response-to-rick-hills/,
}

\item What incentives would states have if the dissent carried the day? Might
they make it harder to multiply parcels (e.g., by enacting barriers to
subdividing land)? For a blog exchange on this point, see
\clause{hills-half-hearted; somin-more-murr}.

\item What do you make of Justice Thomas's dissent? Is he suggesting that the
whole of regulatory takings jurisprudence is without a constitutional basis? The
Rappaport article he cites argues that even if Takings Clause of the Fifth
Amendment does not support regulatory takings jurisprudence, ``there are strong
reasons, based on history, structure, and purpose, to conclude that the Takings
Clause had a different meaning under the Fourteenth Amendment.'' 45 \textsc{San
Diego L. Rev.} at 731.

