\expected{brown-v-voss}

\item What do you make of the reasoning in \textit{Brown}? What is the point of
taking a strict view of the modification of easements if the majority has no
intention of following it through to its logical consequence? Would it be better
to be more flexible about easement law? Or is the dissent's approach preferable?


\item It turns out that the facts in \textit{Brown} were a good deal less
straightforward than the majority indicates. Elizabeth J.
Samuels, \textit{Stories Out of School: Teaching the Case of Brown v. Voss},  16
\textsc{Cardozo L. Rev}. 1445, 1451 (1995) (``What I discovered from the record
in the case, related land and other court records, and interviews with the
parties was predictable in some ways and startling in others. As the reader of
the opinion suspects, the conflict between the neighbors had a dark and bitter
emotional history. As one cannot easily suspect, the physical aspects of the
property differ in legally significant ways from the court's description and
understanding. When the supreme court of Washington decided the case, the
controversy, unbeknownst to the court, was moot. And the party who appears to be
the loser in the opinion was in reality the winner.'' (footnotes omitted)). 

\expected{jacque-v-steenberg}

\item Can you reconcile this case with \textit{Jacque v. Steenberg Homes}?


\item Note the range of possible outcomes in easement disputes. It is not so
simple as saying the property owner has the right to block the would-be easement
holder or not. Cases like \textit{Brown} honor the property interests of the
landowner, but by requiring the payment of damages rather than granting an
injunction. In other words, the trespasser gets to choose whether or not to
continue using the land. Likewise, the right-of-way statutes discussed above
allow passage with payment. Recall similar remedies in the adverse possession
setting in cases of innocent encroachments. What do you think motivates courts
to elect payment remedies in these situations? Why do you think they are not the
norm in property law? 

