Because easements are interests in land, \term[express easement]{express
easements} are subject to the
Statute of Frauds. Failures to comply with the statute may still be enforced in
cases of reasonable detrimental reliance. \textit{See, e.g.},
\textsc{Restatement (Third) of Property (Servitudes)} \S~2.9. 

\paragraph{Third parties}
Easements are often created as part of the transfer of land (e.g., selling a
property, but retaining the right to use its parking lot). Traditionally,
grantors could reserve an easement in the conveyed land for themselves, but
could not create an easement for the benefit of a third party. This rule led to
extra transactions. Where the traditional rule applied, if A wanted to convey to
B while creating an easement for C, A could convey to C who would then convey to
B, while reserving an easement. 

The modern trend discards this restriction. \textit{See, e.g.}, \emph{Minton v.
Long}, 19 S.W.3d 231, 238 (Tenn. Ct. App. 1999). The California Supreme Court
explained:
\begin{quotation}
The rule derives from the common law notions of reservations from a grant and
was based on feudal considerations. A reservation allows a grantor's whole
interest in the property to pass to the grantee, but revests a newly created
interest in the grantor. While a reservation could theoretically vest an
interest in a third party, the early common law courts vigorously rejected this
possibility, apparently because they mistrusted and wished to limit conveyance
by deed as a substitute for livery by seisin. Insofar as this mistrust was the
foundation of the rule, it is clearly an inapposite feudal shackle today.
Consequently, several commentators have attacked the rule as groundless and have
called for its abolition. 

California early adhered to this common law rule. In considering our continued
adherence to it, we must realize that our courts no longer feel constricted by
feudal forms of conveyancing. Rather, our primary objective in construing a
conveyance is to try to give effect to the intent of the grantor. In general,
therefore, grants are to be interested in the same way as other contracts and
not according to rigid feudal standards. The common law rule conflicts with the
modern approach to construing deeds because it can frustrate the grantor's
intent. Moreover, it produces an inequitable result because the original grantee
has presumably paid a reduced price for title to the encumbered property.
\end{quotation}
\emph{Willard v. First Church of Christ, Scientist}, 7 Cal. 3d 473, 476-77
(1972) (citations omitted). The modern \textsc{Restatement} likewise dispenses
with the traditional approach, allowing the direct creation of easements on
behalf of third parties. \textsc{Restatement} \S~2.6. Some jurisdictions
nonetheless retain the bar, citing reliance interests and the prospect that such
easements create instability in title records. \emph{Estate of Thomson v. Wade},
509 N.E.2d 309, 310 (N.Y. 1987) (``The overriding considerations of the public
policy favoring certainty in title to real property, both to protect bona fide
purchasers and to avoid conflicts of ownership, which may engender needless
litigation, persuade us to decline to depart from our settled rule.'' (internal
citation and quotation omitted)). There is an argument that the extra
transactions required by the traditional rule promote better title indexing. The
\textsc{Restatement} observes:
\begin{quote}
To avoid the prohibition, two conveyances must be used: the first conveys the
easement to the intended beneficiary; the second conveys the servient estate to
the intended transferee. The only virtue of the rule is that it tends to ensure
that a recorded easement will be properly indexed in the land-records system,
but there are so many exceptions to the rule, where it is still in force, that
it does not fill that function very well.
\end{quote}
\textsc{Restatement (Third) of Property (Servitudes)} \S~2.6 cmt. a (2000).
\having{hartig-v-stratman}{In light of \mref{hartig-v-stratman}, are}{Are}{Are}
you persuaded that the benefits of a separate transaction for recording purposes
outweigh the costs?

