\expected{narrative-leases-exit-tenant}

\item Landlord leases property to T1 from January 1, 2015 to December 31, 2015. 
On March 1, T1 sold T2 her remaining interest in the property.  On October 1,
T2 rented the property to T3 for two months.  Describe the privity
relationships between all of the parties.  If T3 stops sending rent payments to
Landlord, whom can the Landlord sue to recover the money? 


\item Alger, a landlord, rents a commercial building to Brown for 5 years.  Six
months into the lease, Brown subleases his interest to Clancy for 3 years. 
Clancy then turns around and assigns his interest to Dahl.  Describe the
privity relationships between all of the parties.  If Dahl stops sending rent
checks to Alger, whom can Alger sue to recover the money.


\item Picasso, a landlord, rents an apartment to Renoir for one year.  The lease
contains a provision allowing the tenant to renew the leasehold for a second
year on the same terms.  Renoir assigns his interest in the lease to Seurat. 
Seurat then assigns his interest to Turner.  What are the privity relationships
between the parties? Can Turner exercise the renewal clause in the original
lease? \textit{See} \textit{Castle v. Double Time, Inc.}, 737 P.2d 900 (Okla.
1987) (discussing renewal clauses). 


\item Landlord leases a unit to T1 for ten years beginning in 2010.  In 2012, T1
transfers all of his right to T2 ``for a period of five'' years.  In 2013, T2
subleases to T3 for one year.  What are the privity relationships and whom can
the landlord sue if T3 stops paying rent?


\item L leases a commercial property to T1 for ten years beginning in 2010.  In
2012, T1 assigns all of her interest to T2.  A year later, T2 assigns all of
her interest to T3.  In 2014, T3 subleases to T4 for a term of four years.  In
the sublease contract, T4 agrees to assume ``all of the covenants and
promises'' in the original lease between L and T1.  In 2015, T4's business
fails and she ceases making paying rent.  What are the privity relationships? 
Whom can L sue to recover the unpaid rent money?  

