The following rental agreement is modeled on an actual lease that a friend of
the casebook authors was asked to sign.  Do you see any potential problems for
a tenant?  Would you sign this lease?

\readinghead{Residential Rental Agreement and Contract}

THIS AGREEMENT (hereinafter known as the ``Lease'' or the ``Agreement'') is made
and entered into this 1st day of September 2015, between
\textbf{Peter Rabbit} (hereinafter referred to as the ``Tenant'') and
\textbf{Mr. McGregor} (hereinafter referred to as the ``Landlord'').  In
exchange for valuable consideration, the landlord and tenant agree to the
following:

1.\quad Property.\quad The landlord owns certain real property and improvements
at \textbf{123 Vegetable Garden Way}, \textbf{Potterville, Beatrixia}
(hereinafter referred to as the ``Property'' or the ``Premise'').  The Landlord
wishes to lease the Premise to the Tenant upon the terms and conditions stated
in this Lease.  The Tenant wishes to lease the Premise from the Landlord upon
the terms and conditions stated in this Lease.

2.\quad Term.\quad This agreement shall commence on September 1, 2015 and shall
commence on August 31, 2018 at 11:59 PM.  Upon any termination of the Agreement,
the Tenant will pay off all outstanding bills, remove all personal property from
the Premise, bring the leased premise back to the condition it was in upon
move-in (excepting normal wear and tear), peacefully vacate the premise, return
all keys to the Landlord, and give the Landlord a forwarding address.  

3.\quad Holdovers.\quad If the Tenant holds over after the termination of the lease, a
new tenancy from month-to-month shall be created.  Under the new month-to-month
lease the Tenant shall be responsible for double the agreed upon rent.  

4.\quad Rent.\quad The Tenant shall pay the landlord \$1000 per month as rent for the
entire term of the agreement.  The rent shall be due on the
1\textsuperscript{st} day of each calendar month.  Weekends, holidays, and
religious observances do not excuse the Tenant's obligation to make timely
payments.

5.\quad Delivery of Possession.\quad The Landlord shall not be held liable for any
failure to deliver possession of the Premise by the starting date of the agreed
upon term.

6.\quad Late Fees.\quad A late fee of 5\% shall be due if the rent is received after the
5\textsuperscript{th} day of the month.  A late of 10\% shall be due if the
rent is received after the 10\textsuperscript{th} day of the month.  Acceptance
of a late fee does not affect or waive any other right or remedy the Landlord
may exercise for Tenant's failure to timely pay rent.  

7.\quad Returned Checks.\quad In the event that any payment by the Tenant is returned
for insufficient funds or if the Tenant stops payment, the Tenant will pay
\$100 to the Landlord for each such event, in addition to the Late Fees
described above.  

8.\quad Security Deposit.\quad The Tenant shall deposit with the Landlord \$1500 as a
security deposit for this Agreement.  All interest that accrues on such a
security deposit shall belong to the Landlord alone. The Landlord may use the
deposit money for any and all purposes allowed by law.

9.\quad Utilities.\quad It is the responsibility of the Tenant to obtain all utilities
for the leased Property.  Tenant's failure to make any payment for the
utilities shall constitute a material breach of the agreement.  The Landlord
shall not be held liable for any failure to deliver any utility service or for
any damage caused by a problem with any utility service, whatever the cause of
such problem.  The Tenants do hereby waive any claim for damages that result
from any problem with utility service.

10.\quad Keys.\quad The Tenant shall not install any new locks anywhere on the property
or make any copies of the keys.  The Tenant also shall refrain from providing
any keys to any person not listed on this Agreement.  When the lease
terminates, the Tenant shall return all keys to the Landlord.

11.\quad Pets.\quad No pets of any kind, type, or breed shall be allowed on the property
without the Landlord's express written consent.  This consent, if given, will
require an additional pet deposit.

12.\quad Use of the Premise.\quad The premise shall be used and occupied solely by the
Tenant.  Tenant shall not allow any other person to use or occupy the premise
without first obtaining Landlord's written consent.  No part of the Premise
shall be used at any time during the term for any business, trade, or other
commercial purpose.  Additionally, the tenant agrees to comply with all local,
state, and federal laws, regulations, and ordinances.  No part of the property
may be used in any way that aids or advances a criminal enterprise.

13.\quad Assignments and Subletting.\quad The Tenant shall not license, assign, or sublet
the Property and/or this agreement without the written consent of the Landlord.
 An assignment, subletting or license without the Landlord's written consent
shall be considered absolutely null and void and, at the Landlord's option,
terminate this Agreement.

14.\quad Alterations.\quad The Tenant shall make no alterations to the Premise without
written consent of the Landlord.  If the Tenant makes any unauthorized
improvement, modification, or change to the Property, the landlord has the
option to charge the Tenant the cost of restoring the Premise to its original
condition.  In the event that the Landlord approves an alteration made by the
Tenant, such alternations shall become the property of the Landlord and remain
on the Property.

15.\quad Maintenance \& Repair.\quad Except for normal wear and tear, the Tenant shall
maintain the Premise in the condition it was upon the starting date of the
Agreement.  Should any damages, malfunctions, breakages, or other problems
occur during the course of the Lease, the Landlord shall have a reasonable
amount of time to complete such repairs.  During that time, the Tenant's rent
shall remain due in full and on time despite any hardships such repairs or
delays may cause.  Tenant also has a contractual duty to (1) notify Landlord of
any problems with the leased premise, (2) Deposit all trash, rubbish, refuse,
and garbage in the trash cans provided by the city, (3) keep all windows,
doors, and locks in good order, (4) inspect the fire alarms each and every
month. 

16.\quad Noise.\quad The Tenant and the Tenant's guests shall at all times keep the
level of sound down to a level that does not annoy or interfere with other
residents or neighbors.  

17.\quad Sale of the Property.\quad The Landlord shall have the right to sell or transfer
his ownership of the Property and this Agreement at any time and without
restriction.  Upon sale or transfer of the Landlord's interest, this agreement
may be terminated by either the Landlord or the party who purchases the
Landlord's interest.  The Tenant agrees to release, waive, and hold harmless
the Landlord and the Landlord's successor from all liabilty if such a transfer
occurs. 

18.\quad Access.\quad The Landlord and his agents shall have the right to enter the
Property without notice to inspect the property, make repairs, or show the
property to prospective tenants or purchasers.

19.\quad Condition of the Premise.\quad The Landlord makes no guarantees or warranties
about the condition of the leased premise.  The Tenant assumes all risk of
injury or harm stemming from any accidents or criminal acts occurring on or
around the Premise.  The Tenant agrees to hold the Landlord harmless for all
liability stemming any injury or harm to the Tenant, Tenant's property, or
Tenant's guests.  The Tenant further agrees to indemnify, defend, and hold
harmless the Landlord from any and all claims over the condition of the
premise.  Should the Tenant damage the Premise, he shall indemnify the Landlord
for all costs of repair or replacement within 30 days.

20.\quad Natural Disaster.\quad In the event of a natural disaster, fire, or other
catastrophic event, the Landlord may choose not to repair the Premise, in which
case the Lease shall terminate.  The Landlord may also elect to fix the
Premise, in which case the Tenant must continue to pay the full monthly rent so
long as the repairs are completed within a reasonable time.  In either case,
any and all damages and injuries connected to acts of the Tenant, his guests,
or property shall be the sole financial responsibility of the Tenant.

21.\quad Eminent Domain.\quad If a government or private entity takes the Premise or any
part of the Premise by eminent domain, this Lease shall terminate.  The new
termination shall be the date of the final taking order.  Any award or court
judgment in favor of the Landlord in an eminent domain case or any settlement
award stemming from an eminent domain proceeding shall belong to the Landlord
in full.  The Tenant shall have no claim over such awards. 

22.\quad Attorney's Fees.\quad Tenant agrees to pay all reasonable attorney's fees,
court costs, and other expenses if it becomes necessary for the Landlord to
enforce any of the conditions of covenants of this Lease, including but not
limited to eviction proceedings, collection of rents, and damage to the Premise
caused by the Tenant.  The Tenant also agrees to indemnify the Landlord for all
attorney's fees, court costs, and other expenses that the Landlord may incur
while successfully defending a lawsuit brought by the Tenant.

22.\quad Abandonment.\quad If at any time during the term of this Lease the Tenant
abandons the Premise, the Landlord may obtain possession of the Premise in any
manner provided for by law.  Any personal property left behind shall be
considered abandoned. The Landlord may dispose of such personal property in any
manner he deems fit and is released of all liability for doing so. 

23.\quad Severability.\quad If any portion of this Lease shall be found unenforceable,
invalid, or void under any law or public policy, that portion of the Lease
shall be severed from the remainder of the Agreement.  All remaining portions
of the Agreement will remain in effect and enforceable.

24.\quad Governing Law.\quad This lease shall be governed and interpreted under the laws
of the Commonwealth of Beatrixia.

25.\quad Non-Waiver.\quad No delay or non-enforcement of any term of this Agreement by
the Landlord shall not be deemed a waiver.  All terms and conditions of this
Agreement shall remain fully enforceable should the Landlord seek to enforce
any condition or covenant at a later date, even if the Landlord has
intentionally or unintentionally neglected to do so in a previous instance.  

26.\quad Notices.\quad Any notice required or permitted under this Agreement must be
written on 8{\textonehalf} x 11 paper and sent by United Parcel Service (UPS). 
Notice shall be sent to the address of the Property for the Tenant or to
\textbf{345 Bunny Pie Lane, Potterville, Beatrixia} for the Landlord.

27.\quad Spelling and Grammar.\quad Any mistakes in spelling, grammar, punctuation, or
gender usage shall not be fatal to the Agreement.  Rather, they shall be
interpreted to carry out the intent of the parties. 

28.\quad Default.\quad Tenant shall be in default of this Agreement if he fails to
comply with any covenant, condition or term and/or fails to pay rent when due
and/or causes damage to the Premise during the term which cumulatively equals
or exceeds \$100.  Should the Tenant ever default, the Landlord may
\textbf{with or without notice} either (1) terminate the Lease or (2) terminate
the Tenant's right to possession of the Premise while leaving this Agreement
operative.  If the Landlord elects option (2), the Landlord will have the
immediate right to possess the Premises and the Tenant shall lose all
possessory rights and have the obligation to immediately vacate the Premise. 
However, the Tenant shall still have the duty to pay all rents, fees and
expenses mandated under this Agreement and/or by the judicial system until
either the agreed upon term concludes or the property is re-rented at a monthly
rate not less than the amount owned under this Agreement with any negative
balance owed by the Tenant.  

\begin{tabular}{ll}
\strut\leaders\hrule\hskip 20em\hbox{} & \hbox{}\leaders\hrule\hskip
8em\hbox{} \\
Tenant Signature  & Date \\
\strut\leaders\hrule\hskip 20em\hbox{} & \hbox{}\leaders\hrule\hskip
8em\hbox{} \\
Landlord Signature  & Date \\
\end{tabular}



