\reading{An Exercise in Advertising}

Imagine that you are a lawyer for a newspaper in a large metropolitan area.  The
local chapter of the ACLU has raised concerns that some advertisements in the
classifieds section of your paper violate the Fair Housing
Act.\footnote{Would any of these ads violate the Civil Rights Act of
1866?}  Your boss has asked you to review the ads in
Figure~\ref{f:leases-hypo-advertisements}
for any offending language. Which of the following would you feel comfortable
printing?\footnote{The government does provide some guidance to landlords
worried about triggering FHA liability through their advertisements.  There are,
for example, published lists of ``words to avoid'' and ``acceptable language.''
Although context is important, landlords can generally use these phrases: good
neighborhood, secluded setting, single family home, quality construction, near
public transportation, near places of worship, and assistance animals only.} 

\begin{figure}
\begin{center}
\begin{tabular}{p{0.4\textwidth}|p{0.4\textwidth}}
\multicolumn{1}{c|}{\textbf{For Rent}} & \multicolumn{1}{c}{\textbf{For Rent}}\\
Seeking tenant for 1 bed apt. \$500/m. I only rent to black people. &
New apartment building. \$650/m. Walking distance to synagogue. Great
amenities.\\[6pt]
\hline
\multicolumn{1}{c|}{\textbf{For Rent}} & \multicolumn{1}{c}{\textbf{For Rent}}\\
Great deal! Apt. in exclusive Danbury area. Very selective. Contact ASAP.
\$700/m &
Perfect apt. for rent. Near park. \$400/m. Absolutely no pets. \\[6pt]
\hline
\multicolumn{1}{c|}{\textbf{For Rent}} & \multicolumn{1}{c}{\textbf{For Rent}}\\
Snazzy digs near downtown! Looking for muscular football players to rent rooms.
500/m &
Looking for tenants. Absolutely no lawyers. Only couples. Must show income 3x
monthly rent. \\[6pt]
\hline
\multicolumn{1}{c|}{\textbf{For Rent}} & \multicolumn{1}{c}{\textbf{For Rent}}\\
Seeking new tenants for 2 br. Pref for women---I'm female \& want female
tenants! & Great place by University. \$600/m. Kids ok, but must pay 2x security
dep. Kids = trouble \\
\end{tabular}
\end{center}
\caption{Hypothetical classified advertisements.}
\label{f:leases-hypo-advertisements}
\end{figure}

What about this ad for a roommate on Craigslist?  Is it objectionable to you?
Does it violate the FHA?  Does it matter that the poster is looking for a
\textit{roommate}? Would your answers change if the advertisement read, ``Have
a room available for an able-bodied white man with no children?'' 

\begin{quote}
\edfont
\textbf{\$650 / 1010 sq. ft. -- Have a room available for female Christian
elderly woman none smoker}

\nobreak

Nice comfortable furnished room available access to patio, backyard, 1/2 bath \&
full bath available, wyfi available, cable TV available, This is all for that
special Christian elderly none smoking or drinking church going female. Thank
you

\nobreak

Serious inquiries only!!!
\end{quote}

%\heregraphic[height=6em]{leases-05}

