We may well conclude that certain types of resources should be subject to
private ownership, and we may further conclude that such ownership ought to
entail particular rights of owners. But this would not be sufficient to
establish a system of property rights. We would still need to decide which
things are owned by whom. Certainly, if one of the rights of owners is the right
to alienate, then once something is legitimately owned by someone, that person
can transfer rightful ownership to someone else. (We will study how such
transfers can come about later in this book---indeed we will find that some
transfers can confer the rights of ownership on a transferee even where the
transferor's rights are not so clear-cut. We will also see that there are ways
for things owned by one person to become owned by another person \textit{other
than} by voluntary transfer.)

But even assuming a current owner could trace their rights of ownership back
through a series of successive voluntary transfers from rightful owners---a
\textit{chain of title}, as we will come to call it---the first link in that
chain \textit{must} be something other than a transfer from a prior rightful
owner. What could this something be? How do things go from being unowned to
being owned? Why might we recognize some rules for such initial allocations of
resources over the available alternatives? 

In this chapter, we will examine the most common justification for protecting
someone's rights of ownership: \textit{possession}. The common law holds that
initial ownership of a heretofore unowned thing goes to the \textit{first to
possess} that thing---that first in time is first in right. But as we will see,
this rule is not as straightforward as it may seem. To begin with, reasonable
people may differ as to what constitutes ``possession,'' or what it means to be
``first.'' Our first few cases illustrating this problem deal with first
possession of \textit{chattels} (sometimes called ``personal property'' or
``personalty''---basically any ownable thing that isn't land or attached to
land).

