\reading{John Locke, \textit{Second Treatise of Civil Government}}
\readingcite{Ch. 5 (1690)}

Though the earth, and all inferior creatures, be common to all men, yet every
man has a property in his own person: this no body has any right to but himself.
The labour of his body, and the work of his hands, we may say, are properly his.
Whatsoever then he removes out of the state that nature hath provided, and left
it in, he hath mixed his labour with, and joined to it something that is his
own, and thereby makes it his property. It being by him removed from the common
state nature hath placed it in, it hath by this labour something annexed to it,
that excludes the common right of other men: for this labour being the
unquestionable property of the labourer, no man but he can have a right to what
that is once joined to, at least where there is enough, and as good, left in
common for others.

He that is nourished by the acorns he picked up under an oak, or the apples he
gathered from the trees in the wood, has certainly appropriated them to himself.
No body can deny but the nourishment is his. I ask then, when did they begin to
be his? when he digested? or when he eat? or when he boiled? or when he brought
them home? or when he picked them up? and it is plain, if the first gathering
made them not his, nothing else could. That labour put a distinction between
them and common: that added something to them more than nature, the common
mother of all, had done; and so they became his private right. And will any one
say, he had no right to those acorns or apples, he thus appropriated, because he
had not the consent of all mankind to make them his? Was it a robbery thus to
assume to himself what belonged to all in common? If such a consent as that was
necessary, man had starved, notwithstanding the plenty God had given him\ldots .

But the chief matter of property being now not the fruits of the earth, and the
beasts that subsist on it, but the earth itself; as that which takes in and
carries with it all the rest; I think it is plain, that property in that too is
acquired as the former. As much land as a man tills, plants, improves,
cultivates, and can use the product of, so much is his property. He by his
labour does, as it were, inclose it from the common. Nor will it invalidate his
right, to say every body else has an equal title to it; and therefore he cannot
appropriate, he cannot inclose, without the consent of all his fellow-commoners,
all mankind. God, when he gave the world in common to all mankind, commanded man
also to labour, and the penury of his condition required it of him. God and his
reason commanded him to subdue the earth, i.e. improve it for the benefit of
life, and therein lay out something upon it that was his own, his labour. He
that in obedience to this command of God, subdued, tilled and sowed any part of
it, thereby annexed to it something that was his property, which another had no
title to, nor could without injury take from him.

Nor was this appropriation of any parcel of land, by improving it, any prejudice
to any other man, since there was still enough, and as good left; and more than
the yet unprovided could use\ldots .

God gave the world to men in common; but since he gave it them for their
benefit, and the greatest conveniencies of life they were capable to draw from
it, it cannot be supposed he meant it should always remain common and
uncultivated. He gave it to the use of the industrious and rational, (and labour
was to be his title to it;) not to the fancy or covetousness of the quarrelsome
and contentious. He that had as good left for his improvement, as was already
taken up, needed not complain, ought not to meddle with what was already
improved by another's labour: if he did, it is plain he desired the benefit of
another's pains, which he had no right to, and not the ground which God had
given him in common with others to labour on, and whereof there was as good
left, as that already possessed, and more than he knew what to do with, or his
industry could reach to.

\ldots To which let me add, that he who appropriates land to himself by his
labour, does not lessen, but increase the common stock of mankind: for the
provisions serving to the support of human life, produced by one acre of
inclosed and cultivated land, are (to speak much within compass) ten times more
than those which are yielded by an acre of land of an equal richness lying waste
in common. And therefore he that incloses land, and has a greater plenty of the
conveniencies of life from ten acres, than he could have from an hundred left to
nature, may truly be said to give ninety acres to mankind: for his labour now
supplies him with provisions out of ten acres, which were but the product of an
hundred lying in common. I have here rated the improved land very low, in making
its product but as ten to one, when it is much nearer an hundred to one: for I
ask, whether in the wild woods and uncultivated waste of America, left to
nature, without any improvement, tillage or husbandry, a thousand acres yield
the needy and wretched inhabitants as many conveniencies of life, as ten acres
of equally fertile land do in Devonshire, where they are well cultivated?

\ldots Thus labour, in the beginning, gave a right of property, wherever any one
was pleased to employ it upon what was common, which remained a long while the
far greater part, and is yet more than mankind makes use of. Men, at first, for
the most part, contented themselves with what unassisted nature offered to their
necessities: and though afterwards, in some parts of the world, (where the
increase of people and stock, with the use of money, had made land scarce, and
so of some value) the several communities settled the bounds of their distinct
territories, and by laws within themselves regulated the properties of the
private men of their society, and so, by compact and agreement, settled the
property which labour and industry began; and the leagues that have been made
between several states and kingdoms, either expresly or tacitly disowning all
claim and right to the land in the others possession, have, by common consent,
given up their pretences to their natural common right, which originally they
had to those countries, and so have, by positive agreement, settled a property
amongst themselves, in distinct parts and parcels of the earth\ldots

\ldots Thus in the beginning all the world was America\ldots .

