\editfromrepo{base}


\replacerange{It lies, roughly, in the form}{two railroads are substantially
parallel.}{\ldots}

\replacerange{The use districts are classified in respect}{bill boards, sign
boards and advertising signs\ldots.}{[The opinion lists the permissible uses for
each of the districts in detail. It is sufficient for our purposes that the use
districts form a spectrum from low-density residential (U-1) to heavy industry
(U-6). The height and area districts imposed requirements on the height of
buildings and square footage of lots, respectively.]}

\replacerange{; that it is being enforced in accordance with its
terms}{injunction upon this record in respect of this class of
restrictions.\ldots}{\ldots.}

\replace{problems have developed, and constantly are developing, which require,
and will continue to require, additional restrictions}{problems are
developing, which require additional restrictions}

\append{occupation of private lands in urban communities.}{\edfootnote{Many
of the opinion's sentences are extraordinarily long, and have been edited down
substantially here. Consult the original before quoting anything.}}

\replacerange{Regulations the wisdom, necessity and validity of which}{have been
rejected as arbitrary and oppressive}{Regulations that are now uniformly
sustained probably would have been rejected as arbitrary and oppressive even a
half century ago.}

\replacerange{ In a changing world, it is}{Constitution of course must
fall.}{\ldots}

\replacerange{fixing the height of buildings within reasonable limits}{must be
left open, in order}{[on building height and construction]}

\replacerange{ Such laws may also find their justification in the fact
that}{trade structures, presently to be discussed.}{\ldots}

\replacerange{It is said that the Village of Euclid}{within the village
borders.}{[The real estate owner argued that Cleveland's industrial development
was spreading into suburbs like Euclid, which would increase property values. So
Euclid's zoning ordinances prohibiting industrial uses meant that the owner
would not enjoy those property value increases.]}

\replacerange{ the provisions of the ordinance excluding from residential
districts}{recent zoning legislation, namely,}{\ldots}

\delete{including those of the appellee, or to particular conditions, or to be
considered in connection with specific complaints, }

\deleterange{, as it is here, not upon the ground of}{in process of actual
execution, but}

\delete{, by materially and adversely affecting values and curtailing the
opportunities of the market,}

\deleterange{, to ascertain by a process of piecemeal dissection}{not withstand
the test of constitutionality}

\endedit
