\editfromrepo{base}

\deleterange{\captionedgraphic}{\par}

\replacerange{Justice Marshall invoked both the}{large manor house next
door.'' A}{In Justice Marshall's view, a}

\replacerange{more limited definition of ``family,'' over several dissents.}{In
that case, Inez Moore}{more limited definition of ``family,'' over several
dissents.\footnote{The relevant ordinance defined ``family'' as ``a number of
individuals related to the nominal head of the household or to the spouse''
thereof, and further enumerated specific relatives that could qualify as part of
the statutorily defined family.} In that case, Inez Moore}

\deleterange{ Such ``intrusive regulation of the family''}{and
underinclusiveness.}

\deleterange{There was a long tradition of}{apparently what happened here.'' }

\deleterange{Justice Brennan's concurrence also discussed the
possibility}{authorities choose to prosecute.'' \par}

\deleterange{It was acceptable for a city to choose}{commission were
African-American. }

\replacerange{Black Jack, Missouri, another St. Louis suburb.}{\par}{Black Jack,
Missouri, another St. Louis suburb.\footnote{\textit{See, e.g.}, Nancy Larson,
\textit{Gay Couples Keep 
Out!}, \textsc{Advocate} 34 (Jul. 18, 2006) (discussing lesbian couple and  
daughter who were warned by real estate agents that Ladue would prevent them
from living together); Eun Kyung Kim, \textit{Law Means Unwed Couple, 3 Kids May
Be\ldots Booted From Black Jack}, \textsc{St. Louis Post-Dispatch} A1 (Feb. 22,
2006).}\par}

\replacerange{Its ordinance now defines family as:}{Now that the lesbian
couple}{\par Now that the lesbian couple}

\deleterange{\item Compare Ferguson's definition of family:}{How would you frame
it?}

\replacerange{from living together); }{The Court of Appeals of New York}{from
living together). The Court of Appeals of New York}

\replacerange{to serve no valid purpose'');}{Many New York municipalities}{to
serve no valid purpose''). Many New York municipalities}

\replacerange{if the owner occupied the premises);}{Some jurisdictions have even
tightened}{if the owner occupied the premises). Some jurisdictions have even
tightened}

\replacerange{\item As for college students, can measures}{The FHA and the
ADA}{\item
\textbf{The Fair Housing Act and the Americans with Disabilities Act}. The FHA
and the ADA}




\endedit
