\item Amara and Babar are an unmarried couple who would like to buy a house
together. They want the protections available to tenants by the entirety. How
would you write the deed to create an estate that was as much like tenancy by
the entirety as is legally possible? Would you advise them to do this, or
would you advise them to do something else? If so, what?

\item Abbie and Jenny Mills are sisters who wish to share all their property.
Each wants to retain the ability to leave with a fair share of the assets if
their relationship breaks down or either of them forms a pair bond. Jenny, who
has been unable to find a long-term job after her graduation from college,
agrees to take care of housekeeping, while Abbie agrees to work as an FBI agent
in order to bring in money. Jenny knows that, if this arrangement persists for a
long time, it may be even harder for her to get a job later, because she won't
have much work history. Abbie wants to make sure that, if she needs to move
for work, she will be able to sell their shared house and buy another. Other
than their family home and Abbie's retirement account, they have very few
assets. Draft an agreement between them. What is reasonable to protect
Jenny's interests? What is reasonable to protect Abbie's interests?

\item Wanda and Pietro own equal shares in a house as tenants in common, but
Wanda is traveling the world while Pietro lives in the house, whose fair rental
value is \$2500 per month. Wanda refuses to contribute to the mortgage, property
taxes, and insurance. If Pietro sues Wanda for an accounting and contribution to
these costs, what will be the result if the carrying costs are \$1000 per month?
What if they're \$3000 per month?

Suppose Pietro adds a screened porch to the house at a cost of \$10,000. Later,
Wanda and Pietro sell the house for \$250,000. How should the proceeds be
divided if the screened porch enhances the value of the property by \$15,000?
What if it only increased the value of the property by \$5000?

