The original modules of \emph{Open Source Property} are copyright
2015--\the\year\ by
Stephen Clowney, James Grimmelmann, Michael Grynberg, Jeremy Sheff, and Rebecca
Tushnet. It may be reused under the terms of the Creative Commons Attribution
NonCommercial 4.0 International license,
\url{https://creativecommons.org/licenses/by-nc/4.0/}. More specifically:
\begin{itemize}
\item   All of the authors' contributions, including the selection and
arrangement of
materials and any original notes and questions, are hereby made available for
reuse under the terms of the CC-BY-NC 4.0 license.
\item   All excerpts from cases and statutes are public-domain and may be freely
reused. All materials in which once-applicable copyrights have expired are
public-domain and may be freely reused.
\item   Where excerpts from books, articles, or photos have been used with
permission, we have noted this in the text, footnotes, or captions. Uses other
than as part of these materials may require separate permission from the
copyright owners, or another justification under copyright law.
\item   All other excerpts of copyrighted materials are used herein under the
fair use provisions of 17 U.S.C. \S~107. We think our interpretation of fair use
is right, but we don’t guarantee it, and your situation may be different if you
make a use other than as part of these materials.
\end{itemize}

The \LaTeX\ version of this book was prepared by Charles Duan, with minor
revisions to the content. To the extent that any such contributions are
copyrightable, they are copyright 2023--\the\year\ Charles Duan, and subject to
the same license terms as above.
