\reading{New York Real Property Law}

\readinghead{\S 258---Short forms of deeds and mortgages.}

The use of the following forms of instruments for the conveyance and mortgage of
real property is lawful, but this section does not prevent or invalidate the
use of other forms: 

\readinghead{Statutory Form A (Individual)}

\readinghead{DEED WITH FULL COVENANTS.}

\def\blankunderline{
    \unskip\space\hbox to 6em{\leaders\hrule\hfill}\space\ignorespaces
}


This indenture, made the \blankunderline day of \blankunderline nineteen hundred  and \blankunderline,
between \blankunderline (insert residence) party of the first part, and
\blankunderline (insert residence) party of the second part, 

Witnesseth, that the party of the first part, in consideration of
\blankunderline
dollars, lawful money of the United States, paid by the party of the second
part, does hereby grant and release unto the party of the second part,
\blankunderline and assigns forever, all \blankunderline (description), together
with the
appurtenances and all the estate and rights of the party of the first part in
and to said premises, 

To have and to hold the premises herein granted unto the party of the second
part, \blankunderline and assigns forever. And said \blankunderline covenants as
follows: 

First. That said \blankunderline is seized of said premises in fee simple, and
has good right to convey the same; 

Second. That the party of the second part shall quietly enjoy the said premises;


Third. That the said premises are free from incumbrances; 

Fourth. That the party of the first part will execute or procure any further
necessary assurance of the title to said premises; 

Fifth. That said \blankunderline will forever warrant the title to said
premises. 

In witness whereof, the party of the first part has hereunto set his hand and
seal the day and year first above written. 

In presence of:

\readinghead{Statutory Form D. (Individual)}

\readinghead{QUITCLAIM DEED.}

This indenture, made the \blankunderline day of \blankunderline nineteen hundred
and \blankunderline between \blankunderline (insert residence), party of the
first part, and \blankunderline (insert residence), party of the second part: 

Witnesseth, that the party of the first part, in consideration of
\blankunderline
dollars, lawful money of the United States, paid by the party of the second
part, does hereby remise, release, and quitclaim unto the party of the second
part, \blankunderline and assigns forever, all (description), together with the
appurtenances and all the estate and rights of the party of the first part in
and to said premises. 

To have and to hold the premises herein granted unto the party of the second
part, \blankunderline and assigns forever. 

In witness whereof, the party of the first part has hereunto set his hand and
seal the day and year first above written. 

In presence of: 

