A review of Ferguson's 1998 planning document, the most recent
available,\footnote{In December 2020, the City of Ferguson put forward a new
plan, \textsc{Our Ferguson 2040} (2020),
\url{https://www.fergusoncity.com/DocumentCenter/View/4897/OurFerguson-2040-Comprehensive-Plan}.}
shows the same embrace of Euclidean zoning as Ladue, but with a different
economic context. \textsc{City of Ferguson Vision 2015 Plan Update} (Aug. 1998),
\url{https://www.fergusoncity.com/DocumentCenter/View/536/2015-Vision-Plan-Comprehensive-Plan-1998-2015-}.
As the document explains, Ferguson is approximately 13 miles northwest of
downtown St. Louis, near the interstate highway system. Ferguson was
incorporated in 1894, and a streetcar line to St. Louis was completed in 1900.
The city grew rapidly after World War II, aided by the rise of cars. The
population peaked in the 1970s, then declined 22.5\% between 1970 and 1990, in
line with the experience of many other St. Louis suburbs. From the document's
introductory materials:
\begin{quote}
Ferguson is one of 92 municipalities in St. Louis County. The County's local
government structure is a confusing mass of small municipalities, school
districts, fire protection districts, isolated pockets of unincorporated lands
and special districts. Many cities are in more than one school district and some
cities are protected by more than one fire protection district. For many years,
there have been discussions about consolidating the City of St. Louis and St.
Louis County and all of its municipalities into a single government entity. Such
a government would serve a population of more than 1.3 million people. While
this might seem desirable in that it would cut down on duplication of services,
the likelihood of this occurring in the near future seems remote.
\end{quote}

Ferguson embraces separation of uses as a goal, along with maintaining or
reducing residential density. In the downtown area, Ferguson would like to
encourage mixed-use development, but not to the extent of disrupting existing
residential neighborhoods. The plan considers suburban residential development
at four single-family houses per acre desirable. 

The planning document noted, however, that residential land use generates little
in the way of taxes, either property taxes or sales taxes. Moreover, Ferguson
considered that rental properties were a problem, because ``some owners of
rental property (particularly absentee owners who do not live in the community)
do not maintain the property as well as many owner-occupied dwellings are
maintained,'' causing health and safety problems. The recommended solution was
to require inspections of rental property for any change in occupancy; the
result contributed to the fine-based scheme of city financing, discussed above.
The plan also recommended taking measures to decrease the number of units that
were rented, but that didn't work.

In addition, because of the general downward economic drift of the area,
Ferguson was confronted with new businesses, which were in need of regulation:
``commercial uses such as pawn shops, check cashing agencies and other
establishments which are associated with communities in decline, should be
closely regulated by the city to prohibit the concentration of such uses in any
one area and\ldots such uses should be prohibited near churches, schools and
residences.'' The current code extensively regulates both the location and the
physical configuration of these businesses.

\captionedgraphic{zoning-img006}{Ferguson Master Zoning Map}

Ferguson has 14 zoning districts, four for single-family houses, one for one- or
two-family houses, one for multiple-family residences, a ``planned residence''
district, a planned ``mixed use'' district, a general commercial district, a
planned commercial district, a downtown core business district, a downtown area
business district, an industrial district, and an airport district. Required
minimum lot sizes vary in the single-family districts from a minimum of 20,000
square feet per family to a minimum of 7,500 square feet. The residential
districts also allow public facilities such as parks and museums, community
gardens, communication towers, group homes, foster care homes, churches, and
family day care homes, as well as other assorted uses (stables in one; bed and
breakfast inns in all but the single-family residential district with the
largest minimum lot size; ``urban agriculture'' on 2 acres or more in the
single-family residential district with the smallest minimum lot size). The
two-family residence district is like the 7,500 square feet-minimum residential
district, but also allows two-family homes on lots of at least 5,000 square
feet. The multiple-family residence district also allows multiple-family homes
on lots of at least 5,000 square feet, as well as state licensed nursing
facilities and residential treatment facilities.

The planned residence district is supposed to mix types of housing, and also
allows adult day care. The planned mixed-use district is similar, with the
addition of ``commercial, cultural, and institutional uses'' and the goals of
minimizing car travel and putting employment and retail closer to higher-density
housing. The allowed commercial activities are limited.\footnote{Private clubs
or lodges; retail sales including appliance, bakery, book store, card and gift
shop, carpeting, clothing, department store, drug store, electronics, fabrics,
food store, furniture store, furrier, garden shop, hardware store, health foods,
hobby shop, ice cream parlor, jewelry store, liquor store, newsstand, pet shop,
radio and T.V. stores and sporting goods; financial institutions without
drive-up facilities, and offices including business, dental, laboratory testing,
medical, research and veterinarian. Special permits are available for
residential treatment facilities; group quarters; billiard parlors, bowling
alleys, racquetball courts, tennis facilities, theaters, restaurants and bars;
drive-through facilities for financial institutions; and Automated Teller
Machines (ATMs).} The general commercial district ``is designed to allow
considerable latitude in the range of retail uses allowed, provided that the
uses are legal and no outdoor storage is conducted'' except as specifically
allowed by the code. What that ``considerable latitude'' looks like can be seen
by skimming (\textit{please} do not try to grasp every detail) the following
list of allowed uses: 
\begin{itemize}
\item Agricultural Services

\item Veterinary Services

\item Transportation and Communication Uses.

\item Local and Suburban Transit and Interurban Highway Passenger Transportation

\item U.S. Postal Service

\item Communication 

\item Building Materials, Hardware, and Garden Supply, including only the
following:
\begin{itemize}
\item Lumber and Other Building Materials Dealers including only:
\begin{itemize}
\item Doors---retail.

\item Fencing dealers---retail.

\item Flooring, wood---retail.

\item Garage doors---retail.

\item Lumber and building material dealers---retail.

\item Lumber and planing mill product dealers---retail.

\item Millwork and lumber dealers---retail.

\item Paneling---retail.

\item Storm windows and sash, wood or metal---retail.
\end{itemize}
\item Paint, Glass, and Wallpaper Stores 

\item Hardware Stores 

\item Retail Nurseries, Lawn and Garden Supply Stores 
\end{itemize}
\item General Merchandise Stores 

\item Food Stores 

\item Automotive Dealers and Gasoline, Service Stations, including only the
following:
\begin{itemize}
\item Motor Vehicle Dealers-New or New and Used.

\item Auto and Home Supply Stores.

\item Gasoline Service Stations.

\item Boat Dealers.

\item Recreational and Utility Trailer Dealers.

\item Motorcycle Dealers.

\item Other New Automotive Dealers.
\end{itemize}
\item Apparel and Accessory Stores.

\item Furniture, Home Furnishings and Equipment Stores.

\item Eating Places (provided that such does not have a drive-through window and
is not a drive-in business).

\item Miscellaneous Retail, including only the following:
\begin{itemize}
\item Drug Stores and Proprietary Stores.

\item Liquor Stores.

\item Miscellaneous Shopping Goods Stores.

\item Retail Stores, Not Elsewhere Classified (except auction rooms); the term
``Retail Stores, Not Elsewhere Classified'' shall not include adult-related
businesses, including adult bookstores, adult novelty shops, and adult retail
stores\ldots.
\end{itemize}
\item Depository Institutions, including only the following:
\begin{itemize}
\item Central Reserve Depository Institutions.

\item Commercial Banks.

\item Savings Institutions.

\item Credit Unions.

\item Foreign Banking.
\end{itemize}
\item Functions Related to Depository Banking, including:
\begin{itemize}
\item Non-deposit Trust Facilities.

\item Functions Related to Depository Banking, Not Elsewhere Classified (except
Check Cashing agencies).
\end{itemize}
\item Non-depository Credit Institutions, including only the following:
\begin{itemize}
\item Federal and Federally-Sponsored Credit Agencies.

\item Business Credit Institutions.

\item Mortgage Bankers and Brokers.

\item Security and Commodity Brokers, Dealers, Exchanges, and Services 
\end{itemize}
\item Insurance

\item Insurance Agents, Brokers, and Service 

\item Real Estate, including only the following:
\begin{itemize}
\item Real Estate Operators (except Developers) and Lessors, including only the
following:

\item Operators of Nonresidential Buildings 

\item Operators of Apartments.

\item Operators of Dwellings Other Than Apartment Buildings.

\item Real Estate Agents and Managers 

\item Title Abstract Offices

\item Sub-dividers and Developers, Except Cemeteries
\end{itemize}
\item Holding and Other Investment Offices 

\item Hotels, Motels, and Tourist Courts 

\item Personal Services, including only the following:
\begin{itemize}
\item Laundry, Cleaning, and Garment Services, including only the following:
\begin{itemize}
\item Garment Pressing and Agents for Laundries and Dry Cleaners 

\item Coin-Operated Laundries and Dry Cleaning.

\item Laundry and Garment Services, not elsewhere classified.
\end{itemize}
\item Photographic Studios, Portrait

\item Beauty Shops 

\item Barber Shops 

\item Shoe Repair Shops and Shoeshine Parlors 

\item Funeral Service and Crematories 

\item Miscellaneous Personal Services.
\end{itemize}
\item Tax Return Preparation Services.

\item Miscellaneous Personal Services, Not Elsewhere Classified (except escort
services, massage parlors, steam baths, tattoo parlors, and Turkish baths).

\item Business Services, including only the following:
\begin{itemize}
\item Advertising 

\item Consumer Credit Reporting Agencies, Mercantile Reporting Agencies, and
Adjustment and Collection Agencies 

\item Mailing, Reproduction, Commercial Art and Photography, and Stenographic Services

\item Miscellaneous Equipment Rental and Leasing, including only the following:
\begin{itemize}
\item Medical Equipment Rental and Leasing 

\item Equipment Rental and Leasing, Not Elsewhere Classified (except airplane
rental and leasing, industrial truck rental and leasing, oil field equipment
rental and leasing, and oil well drilling equipment rental and leasing)
\end{itemize}
\item Personnel Supply Services 

\item Computer Programming, Data Processing, and Other Computer Related Services

\item Miscellaneous Business Services, including only the following:
\begin{itemize}
\item Detective, Guard, and Armored Car Services 

\item Security Systems Services

\item News Syndicates 

\item Photo finishing Laboratories 

\item Business Services, Not Elsewhere Classified (except Gas systems, contract
conversion from manufactured to natural gas, and Scrap steel cuffing on a
contract or fee basis).
\end{itemize}
\end{itemize}
\item Automotive Repair, Services, and Garages, including only the following:
\begin{itemize}
\item Automotive Rental and Leasing, Without Drivers.

\item Automobile Parking.

\item Car Washes (except bus washing and truck washing).
\end{itemize}
\item Miscellaneous Repair Services, including only the following:
\begin{itemize}
\item Electrical Repair Shops.

\item Watch, Clock, and Jewelry Repair.

\item Re-upholstery and Furniture Repair.
\end{itemize}
\item Motion Pictures.

\item Amusement and Recreation Services.

\item Health Services.

\item Legal Services.

\item Educational Services.

\item Social Services, including only the following:
\begin{itemize}
\item Individual and Family Social Services including Adult Day Care Centers.

\item Job Training and Vocational Rehabilitation Services.
\end{itemize}
\item Museums, Art Galleries, Botanical and Zoological Gardens

\item Membership Organizations, including only the following:
\begin{itemize}
\item Business Associations.

\item Professional Membership Organizations.

\item Labor Unions and Similar Labor Organizations.

\item Civic, Social, and Fraternal Associations.

\item Political Organizations.

\item Other Membership Organizations.
\end{itemize}
\item Engineering, Accounting, Research, Management, and Related Services,
including only the following:
\begin{itemize}
\item Engineering, Architectural, and Surveying Services.

\item Accounting, Auditing, and Bookkeeping Services.

\item Research, Development, and Testing Services, including only the following:

\item Commercial Economic, Sociological, and Educational Research.
\end{itemize}
\item Management and Public Relations Services, including only the following:
\begin{itemize}
\item Management Services.

\item Management Consulting Services.

\item Public Relations Services.
\end{itemize}
\item Business Consulting Services, Not Elsewhere Classified.

\item Services Not Elsewhere Classified.

\item Public Administration, including only the following:
\begin{itemize}
\item Executive, Legislative, and General Government, except Finance.

\item Justice, Public Order, and Safety, including only the following:
\begin{itemize}
\item Courts.

\item Police Protection.

\item Legal Counsel and Prosecution.

\item Fire Protection.

\item Other Public Order and Safety.
\end{itemize}
\item Public Finance, Taxation, and Monetary Policy.

\item Administration of Human Resources Programs.

\item Administration of Environmental Quality and Housing Programs.

\item Administration of Economic Programs.

\item National Security and International Affairs.
\end{itemize}
\end{itemize}

Allowed with a permit:
\begin{itemize}
\item Used Vehicle Sales 

\item Used Merchandise Sales and Auction Rooms.

\item Antique stores.

\item Book stores, secondhand.

\item Clothing stores, secondhand.

\item Furniture stores.

\item Furniture, antique.

\item Glassware, antique.

\item Home furnishings, secondhand.

\item Home furnishings, antique.

\item Musical instrument stores, secondhand.

\item Objects of art, antique.

\item Pawnshops

\item Phonograph and phonograph records stores, secondhand.

\item Shoe stores, secondhand.

\item Auction rooms.

\item Check Cashing Agencies and Personal Credit Institutions (except Short-term
Loan Establishments)\footnote{With limits on how close they can be to places of
worship, schools, and residential zones; requirements that each store be at
least 1,000 feet from similar stores including pawnshops; limits on hours of
operations; bans on walk-up or drive-up windows; bans on having bars, heavy mesh
screens or similar material visible from outside; and other restrictions.} 

\item Miscellaneous Personal Services.\footnote{Escort services, massage
parlors, steam baths, tattoo parlors, and Turkish baths may be allowed, subject
to similar restrictions on locations near places of worship, schools,
residentially zoned property, pawn shops, check cashing establishments, and any
other miscellaneous personal service establishment.} 

\item Automotive Repair Shops and Automotive Services, Except Repair 

\item Religious Organizations 

\item General Warehousing and Storage including only Mini-warehouses and
Self-Service Storage Facilities

\item Child Day Care Services 

\item Automated Teller Machines (ATMs).

\item Eating Places (all uses which have drive-through windows or is a drive-in
business) and Drinking Places (Alcoholic Beverages).
\end{itemize}

The planned commercial district is supposed to be planned as a unit, with a
narrower range of allowable commercial uses. The industrial district allows
light manufacturing and wholesale uses:
\begin{itemize}
\item Construction Uses including only the following:
\begin{itemize}
\item Building Construction---General Contractors and Operative Builders.

\item Heavy Construction other than Building Construction---Contractors.

\item Construction---Special Trade Contractors.
\end{itemize}
\item Manufacturing Uses including only the following:
\begin{itemize}
\item Bakery Products.

\item Bottled and Canned Soft Drinks and Carbonated Waters.

\item Manufactured Ice.

\item Textile Mill Products.

\item Apparel and other Finished Products made from Fabrics and Similar
Materials.

\item Millwork, Veneer, Plywood, and Structural Wood Members.

\item Wood Containers.

\item Wood Buildings and Mobile Homes.

\item Miscellaneous Wood Products.

\item Furniture and Fixtures.

\item Paperboard Containers and Boxes.

\item Converted Paper and Paperboard Products, Except Containers and Boxes.

\item Printing, Publishing, and Allied Industries.

\item Rubber and Miscellaneous Plastics Products.

\item Leather and Leather Products.

\item Flat Glass.

\item Glass and Glassware, Pressed or Blown.

\item Glass Products, Made of Purchased Glass.

\item Pottery and Related Products.

\item Metal Cans and Shipping Containers.

\item Cutlery, Hand Tools and General Hardware.

\item Heating Equipment and Plumbing Fixtures.

\item Fabricated Structural Metal Products.

\item Screw Machine Products, and Bolts, Nuts, Screws, Rivets, and Washers.

\item Metal Forgings and Stampings.

\item Coating, Engraving, and Allied Services.

\item Miscellaneous Fabricated Metal Products.

\item Machinery.

\item Electronic and other Electrical Equipment and Components.

\item Measuring, Analyzing, and Controlling Instruments, Photographic, Medical,
and Optical Goods; Watches and Clocks.

\item Miscellaneous Manufacturing Industries.
\end{itemize}
\item Transportation and Communication Uses Including only the following:
\begin{itemize}
\item Local and Suburban Transit and Interurban Highway Passenger
Transportation.

\item Motor Freight Transportation and Warehousing.

\item U.S. Postal Service.

\item Pipe Lines, Except Natural Gas.

\item Transportation Services.

\item Communication.

\item Communication antennae.

\item Communication towers.
\end{itemize}
\item Electric, Gas, and Sanitary Services including only the following:
\begin{itemize}
\item Electric Services, including facilities which are engaged in the
transmission and/or distribution of electric energy for sale.

\item Natural Gas Distribution.

\item Sanitary Services, Not Elsewhere Classified.
\end{itemize}
\item Steam and Air-Conditioning Supply.

\item Irrigation Systems.

\item Wholesale Trade---Durable Goods including only the following:
\begin{itemize}
\item Motor Vehicles and Automotive Parts and Supplies.

\item Furniture and Home Furnishings.

\item Lumber and other Construction Materials.

\item Professional and Commercial Equipment and Supplies.

\item Electrical Goods.

\item Hardware, Plumbing and Heating Equipment and Supplies.

\item Machinery, Equipment, and Supplies.

\item Miscellaneous Durable Goods.
\end{itemize}
\item Wholesale Trade---Nondurable Goods including only the following:
\begin{itemize}
\item Paper and Paper Products.

\item Drugs, Drug Proprietaries and Druggists' Sundries.

\item Apparel, Piece Goods, and Notions.

\item Groceries and Related Products.

\item Beer, Wine, and Distilled Alcoholic Beverages.

\item Books, Periodicals, and Newspapers.

\item Flowers, Nursery Stock, and Florists' Supplies.

\item Tobacco and Tobacco Products.

\item Miscellaneous Nondurable Goods, Not Elsewhere Classified.
\end{itemize}
\item Retail Trade including only the following:
\begin{itemize}
\item Lumber and other Building Materials Dealers.

\item Paint, Glass and Wallpaper Stores.

\item Hardware Stores.

\item Retail Nurseries, Lawn and Garden Supply Stores.

\item Gasoline Service Stations.

\item Furniture, Home Furnishings, and Equipment Stores.
\end{itemize}
\item Services including only the following:
\begin{itemize}
\item Laundry, Cleaning, and Garment Services.

\item Business Services.

\item Automotive Repair, Services and Garages.

\item Miscellaneous Repair Services.
\end{itemize}
\item Amusement and Recreation Services including only the following:
\begin{itemize}
\item Commercial Sports.

\item Miscellaneous Amusement and Recreation Services.
\end{itemize}
\item Health Services including only the following:
\begin{itemize}
\item Medical and Dental Laboratories.

\item Health and Allied Services.
\end{itemize}
\item Educational Services including only the following:
\begin{itemize}
\item Correspondence Schools and Vocational Schools.

\item Research/Development and Testing Services.

\item Miscellaneous Services.
\end{itemize}
\item Public Administration including only the following:
\begin{itemize}
\item Executive, Legislative, and General Government.

\item Police Protection.

\item Fire Protection.
\end{itemize}
\item Adult-Related Business.
\end{itemize}

By permit:
\begin{itemize}
\item Eating Establishments: (all uses, excluding outdoor seating).

\item Utilities:
\begin{itemize}
\item Water Supply storage.

\item Sewage Treatment Plants.
\end{itemize}
\item Adult Entertainment Establishments 

\item Parking facilities for Tractor Trailers 

\item Short-Term Loan Establishments\footnote{Subject to further regulation,
including to avoid ``over-concentration,'' meaning ``a similar use within two
miles of the proposed establishment or more than one such establishment per
10,000 population.'' Where an over-concentration of such uses is found, permit
shall be granted. Distance, hours, and other restrictions apply, including that
the property shall not also be used to issue money orders, cash checks, or sell
lottery tickets; no repossessed property or cars can be stored on site; no extra
advertising materials are allowed, including balloons, lights, flags, etc.; no
writing, printing, or color is allowed on the exterior except for phone number
and office hours in two-inch letters and numbers; and security guards are
required.}
\end{itemize}
The other districts add little to this list, other than an airport.

