\captionedgraphic{zoning-img001}{Houses under construction, Fairfax, VA, by
Zachary Schrag, Aug. 24, 2015}

\term[zoning]{Zoning} is a perennial issue for local governments. For most
homeowners, their
home is their largest asset, and they are exquisitely sensitive to any threats
to its value---but threats can mean either the behavior of their neighbors, or
constraints on their own behavior, setting up a seemingly irresolveable tension.
(Economist William Fischel calls them ``homevoters'' in recognition of the way
that their property interests shape their political choices.) In addition, local
governments and would-be developers of new properties have interests of their
own. Developers too seek to maximize their own property values, including their
ability to develop future projects, which may lead them to sacrifice the
theoretical maximum value of any given parcel. Governments want to protect their
authority and their revenues, goals which they try to accomplish in a variety of
ways.

Zoning is a way of answering the question: What---and where---do we want the
places where we live to be? Our goals in this chapter are to understand the
justifications for and modern varieties of zoning. As you read and review,
consider how zoning compares to other types of land use controls, including
nuisance, private covenants, and the implied warranty of habitability.

Many of our examples in this chapter will come from St. Louis, Missouri, and its
surrounding suburbs. We focus on St. Louis not because it is unique, but because
property law developments in and around St. Louis are broadly representative of
the evolution of metropolitan areas around the country over the past century.
Missouri allows particularly easy formation of new cities from unincorporated
land, and that has contributed to the proliferation of local governments, so
some of the issues are presented particularly starkly in Missouri. Nonetheless,
you should expect similar dynamics to operate throughout the United States.


\captionedgraphic{zoning-img002}{OpenStreetMap map of St. Louis,
\protect\url{https://commons.wikimedia.org/wiki/File:Location_map_USA_St._Louis.png},
BY-SA}


