\expected{planning-zoning-procedures}

\item
Despite the formal insistence on a division between the plan and the
implementation of the plan through zoning, many states allow a zoning ordinance
to be treated as the plan itself. This collapse between planning and zoning was
almost coextensive with the implementation of zoning. The New York City zoning
ordinance imitated by other American cities was, according to Mel Scott, ``a
setback to the city planning movement because it contributed to the widespread
practice of zoning before planning and, in many cities, to the acceptance of
zoning as a substitute for planning.'' \textsc{Mel Scott, American City Planning
Since 1890: A History Commemorating the Fiftieth Anniversary of the American
Institute of Planners} (1971). In the 1920s, three times as many cities adopted
zoning as adopted master plans.

Why is zoning acceptable without a separate master plan? How is a court to judge
its rationality or reasonability without a master plan?

\item
Over time, the tendency in zoning was towards complexity. The city of Euclid had
six use zoning districts in 1926, but now has almost twenty. It wasn't alone:
70\% of municipalities made their zoning rules more restrictive between 1997 and
2002, while only 16\% made them less restrictive. Between 1976 and 2002, the
percentage of zoning decisions that took over two years doubled.

\item
For further background on zoning concepts, New York City's Zoning website,
\url{http://157.188.76.60:8004/site/planning/zoning/about-zoning.page?}, is a
helpful guide.

\item
\textbf{Two examples}. The pages that follow offer descriptions of and portions
from two cities' planning documents and zoning ordinances. Consider the
similarities and differences between Ladue and Ferguson---rather than reading
every word of the ordinances, you should skim them to get a sense of the
behaviors and uses these cities believe they need to regulate. What would each
city do if a new business, say an e-cigarette store (to take a category of
business that did not exist a decade ago) wanted to open?

\expectnext{intro-ladue-mo}

