\reading{Note on Subsequent Purchasers}

\textit{Bevilacqua v. Rodriguez}, 460 Mass. 762, 955 N.E.2d 884 (2011), dealt
with property owners with defective title resulting from
\textit{Ibanez}-style foreclosure problems earlier in the chain. 
In other words, the mortgagee (or, realistically, its representative/putative
representative) had foreclosed in a manner held unlawful by \textit{Ibanez},
then had sold the house again to a new buyer. The Massachusetts high court
held that the new buyers could not clear title under the Massachusetts ``try
title'' procedure, which is a way that an owner can quiet title and establish
which of competing claims is valid. However, the court ruled, that procedure
is only available to people who can plausibly claim to be owners. In
\textit{Bevilacqua}, the chain of title had been broken by the unsuccessful
foreclosure before the purchase, and so the new buyer couldn't bring a
plausible claim. This is not a terribly surprising result: if someone records
a deed to the Brooklyn Bridge, then brings a try title claim to confirm her
ownership, title to the bridge is not conveyed magically even if the true owner
fails to show up. 

The \textit{Bevilacqua} court left open the possibility that owners/lenders
could try to put the chain of title back together and conduct a new, valid
foreclosure, though this will certainly prove complicated in practice, as the
notes above suggest. Another possibility is to track down the old
preforeclosure owner (who is still the owner because of the \textit{Ibanez}
problem) and obtain a quitclaim deed from her. If you represented an old owner
in this situation, what would you counsel? What if you represented a new
buyer?

Is it fair to strip the new buyer of title, when the buyer is unlikely to have
any responsibility (or even much understanding of) the shoddy recording
practices that caused the problem? How else should we resolve the problem? 
Bevilacqua/his title insurer will have a claim against the seller, but we still
need to know who gets the house---you should be able to see similarities from
our discussion of stolen property.

