\item Peter and Nathan Petrelli are the record owners of land in Disturbia. They
hold the land as joint tenants. Peter conveys his interest to Angela Petrelli,
who pays fair market value for it. Angela gives her interest to her
granddaughter Claire Bennet, who records. Peter and Nathan then sell the land
to Matt Parkman, who doesn't know anything about Peter's earlier conveyance to
Angela. Matt records. Disturbia's recording statute reads in relevant part as
follows: ``Every conveyance is void as against any subsequent purchaser of the
same property, or any part thereof, in good faith and for a valuable
consideration, whose conveyance is first duly recorded.'' Who owns what, and
why?


\item Peter Bishop and Olivia Dunham are the record owners of Fringeacre as
joint tenants with right of survivorship. Bishop and Dunham grant Bishop's
father, Walter Bishop, an easement to pasture his cows on Fringeacre. Walter
does not record the easement. ~Later, Dunham gives her interest in Fringeacre
as a gift to Astrid Farnsworth, who does not know about Walter Bishop's
easement, and who records. When Dunham subsequently dies, Peter Bishop sells
Fringeacre to Philip Broyles, who had seen Walter's cows grazing on the
property; when Broyles asked Peter about the cows, Peter said, ``Yeah, I let my
dad do that.'' Broyles records. The relevant recording statute is: ``No
conveyance or mortgage of an interest in land is valid against any subsequent
purchaser for value without notice thereof, unless it is recorded.'' Who owns
what, and why? How, if at all, would the answer change if the recording statute
read, ``Every unrecorded conveyance of an interest in land shall be void as
against any subsequent purchaser in good faith and for a valuable consideration
whose conveyance shall be first duly recorded''?


\item Damon Salvatore conveys Blackacre to Bonnie Bennet. Salvatore subsequently
conveys Blackacre to Caroline Forbes for valuable consideration. Forbes lacks
knowledge of the deed to Bennet. Bennet records, then Forbes records.
Thereafter, Forbes sells the western half of Blackacre to Jeremy Gilbert, and
the eastern half back to Salvatore. Who owns Blackacre and why? Answer with
respect to each of the following laws (Hint: first identify what type of
statute this is): 
\begin{enumerate}
\item {}``No conveyance or mortgage of an interest in land is valid against any
subsequent purchaser for value without notice thereof, whose conveyance is
first recorded.''
\item {}``No conveyance or mortgage of an interest in land is valid against any
subsequent purchaser for value without notice thereof, unless it is recorded.''
\item {}``No conveyance or mortgage of an interest in land is valid against any
subsequent purchaser whose conveyance is first recorded.''
\end{enumerate}

