\item Characterize the following easements using the terms above. Is it
affirmative or negative? Appurtenant or in gross? An easement of access, a
profit, or something else? What is the dominant estate and the servient estate?
Who has the benefit, and who has the burden?

\begin{enumerate}
\item Anna owns Blackacre. Brad has the right to walk across Blackacre to reach
his own property, Whiteacre.

\item Anna owns Blackacre, which has several blackberry bushes. Connie, a chef,
has the right to come onto Blackacre and harvest blackberries.

\item Anna owns Blackacre, which is at the bottom of a hill. Durant owns
Whiteacre, at the top of the hill, with a house on the hill. There is a
retaining wall on Blackacre, preventing the soil from the hill from falling down
(and bringing Durant's house down with it). A court prohibits Anna from tearing
down the retaining wall. (This is called an \term{easement of support}.)
\end{enumerate}
To be sure, there are a lot of terms here, and it is not necessary to fully
memorize them all. But the exercise of categorizing easements and the relevant
parties will help you keep track of the relevant rights and duties, especially
when ownership starts changing hands.

\item The terminology suggests a sharp distinction between affirmative
easements, that give the benefit holder a right to do something, and negative
easements, which impose a duty on the burdened estate not to do something. Is
the distinction really so sharp? In the examples above, doesn't Anna have a duty
not to build a fence to keep Brad out? Doesn't Durant have a right to keep the
retaining wall in place?\having{hohfeld-fundamental}{ (This is the power of the
Hohfeldian theory that rights and duties always come in pairs.)}{}{}

\having{intro-transferring-easements}{}{}{
\item The notion of easements appurtenant ``running with the land'' should tell
you what happens when a dominant estate holder conveys land. If Brad sells
Whiteacre to Ennis, then Ennis can of course use the easement to walk over
Anna's Blackacre. Similarly, the general rule is that the benefit holder
may transfer an easement in gross, absent contrary intent at the time the
easement was created. The burden of an easement obviously transfers with any
conveyance the servient estate. (Why?)
\endgraf
\defcase{brown-v-voss}{
Brown v. Voss, 715 P.2d 514 (Wash. 1986)
}
Things can get much more complicated, though, if the benefit holder seeks to
aggregate or divide up rights. The original \emph{Open Source Property} module
on Easements offers much more detail, but test your intuitions on a few
possibilities (still using the hypotheticals from above):
\begin{itemize}
\item What if Brad also buys Greenacre, and starts using the easement to reach
both Whiteacre and Greenacre? \sentence{see brown-v-voss}.
\item What if Brad builds a 100-unit condominium on Whiteacre? Should all the
unit owners get to use the easement?
\item What if Connie hires three assistants to help her harvest blackberries?
What if Connie's restaurant is bought out by a multinational food chain that now
intends to send out heavy harvesting machinery to Blackacre?
\end{itemize}
}

\item \emph{United States v.~Turoff}, \having{us-v-turoff}{presented earlier in
this book}{presented later in this book}{701 F. Supp. 981 (E.D.N.Y. 1988)},
considers the nature of government-issued taxicab medallions that authorize the
bearer to operate a taxi service:
\begin{quote}
The government also contends that the medallion is, in essence, the equivalent
of an easement to use the city streets. At the risk of dwelling too long on the
esoterica of property, the medallions could not properly be equated with
easements. An easement is generally appurtenant, which is to say that it is a
right which the owner of one parcel of land (the dominant tenement) may
exercise in or over the land of another (the servient tenement) for the benefit
of the former. An easement in gross is a right created in a person to use the
land of another, which the owner of that easement may enjoy even though he does
not own or possess a dominant estate. Although the concept of an easement in
gross has been recognized, such an easement is rare.
\end{quote}
You should now be equipped to follow and evaluate the logic of this argument.
