%
% Abbreviations for court documents, from Indigo Book Table 18
%
\hi@abbrev{Admission}{Admis.}
\hi@abbrev{Admissions}{Admiss.}
\hi@abbrev{Amended}{Am.}
\hi@abbrev{Amendeds}{Ams.}
\hi@abbrev{Appendix}{App.}
\hi@abbrev{Appendixes}{Apps.}
\hi@abbrev{Appendices}{Apps.}
\hi@abbrev{Application}{Appl.}
\hi@abbrev{Applications}{Appls.}
\hi@abbrev{Argument}{Arg.}
\hi@abbrev{Arguments}{Args.}
\hi@abbrev{Attachment}{Attach.}
\hi@abbrev{Attachments}{Attachs.}
\hi@abbrev{Attorney}{Att'y}
\hi@abbrev{Attorneys}{Att'ys}
\hi@abbrev{Brief}{Br.}
\hi@abbrev{Briefs}{Brs.}
\hi@abbrev{Certiorari}{Cert.}
\hi@abbrev{Certioraris}{Certs.}
\hi@abbrev{Civil}{Civ.}
\hi@abbrev{Civils}{Civs.}
\hi@abbrev{Counterclaim}{Countercl.}
\hi@abbrev{Counterclaims}{Countercls.}
\hi@abbrev{Court}{Ct.}
\hi@abbrev{Courts}{Cts.}
\hi@abbrev{Cross-claim}{Cross-cl.}
\hi@abbrev{Cross-claims}{Cross-cls.}
\hi@abbrev{Declaration}{Decl.}
\hi@abbrev{Declarations}{Decls.}
\hi@abbrev{Defendant}{Def.}
\hi@abbrev{Defendants}{Defs.}
\hi@abbrev{Defendant's}{Def.'s}
\hi@abbrev{Defendants}{Defs.}
\hi@abbrev{Defendantses}{Defss.}
\hi@abbrev{Defendants'}{Defs.'}
\hi@abbrev{Demurrer}{Dem.}
\hi@abbrev{Demurrers}{Dems.}
\hi@abbrev{Deny}{Den.}
\hi@abbrev{Denies}{Dens.}
\hi@abbrev{Denying}{Den.}
\hi@abbrev{Denyings}{Dens.}
\hi@abbrev{Deposition}{Dep.}
\hi@abbrev{Depositions}{Deps.}
\hi@abbrev{Discovery}{Disc.}
\hi@abbrev{Discoveries}{Discs.}
\hi@abbrev{Evidence}{Evid.}
\hi@abbrev{Evidences}{Evids.}
\hi@abbrev{Exhibit}{Ex.}
\hi@abbrev{Exhibits}{Exs.}
\hi@abbrev{Hearing}{Hr'g}
\hi@abbrev{Hearings}{Hr'gs}
\hi@abbrev{Injunction}{Inj.}
\hi@abbrev{Injunctions}{Injs.}
\hi@abbrev{Instruction}{Instr.}
\hi@abbrev{Instructions}{Instrs.}
\hi@abbrev{Interrogatory}{Interrog.}
\hi@abbrev{Interrogatories}{Interrogs.}
\hi@abbrev{Joint Appendix}{J.A.}
\hi@abbrev{Joint Appendixes}{J.As.}
\hi@abbrev{Joint Appendices}{J.As.}
\hi@abbrev{Judgment}{J.}
\hi@abbrev{Judgments}{Js.}
\hi@abbrev{Limine}{Lim.}
\hi@abbrev{Limines}{Lims.}
\hi@abbrev{Memorandum}{Mem.}
\hi@abbrev{Memoranda}{Mems.}
\hi@abbrev{Memorandums}{Mems.}
\hi@abbrev{Minute}{Min.}
\hi@abbrev{Minutes}{Mins.}
\hi@abbrev{Motion}{Mot.}
\hi@abbrev{Motions}{Mots.}
\hi@abbrev{Objection}{Obj.}
\hi@abbrev{Objections}{Objs.}
\hi@abbrev{Opinion}{Op.}
\hi@abbrev{Opinions}{Ops.}
\hi@abbrev{Opposition}{Opp'n}
\hi@abbrev{Oppositions}{Opp'ns}
\hi@abbrev{Petition}{Pet.}
\hi@abbrev{Petitions}{Pets.}
\hi@abbrev{Petitioner}{Pet'r}
\hi@abbrev{Petitioners}{Pet'rs}
\hi@abbrev{Petitioner's}{Pet'r's}
\hi@abbrev{Petitioners}{Pet'rs}
\hi@abbrev{Petitionerses}{Pet'rss}
\hi@abbrev{Petitioners'}{Pet'rs'}
\hi@abbrev{Plaintiff}{Pl.}
\hi@abbrev{Plaintiffs}{Pls.}
\hi@abbrev{Plaintiff's}{Pl.'s}
\hi@abbrev{Plaintiffs}{Pls.}
\hi@abbrev{Plaintiffses}{Plss.}
\hi@abbrev{Plaintiffs'}{Pls.'}
\hi@abbrev{Points and Authorities}{P. \& A.}
\hi@abbrev{Preliminary}{Prelim.}
\hi@abbrev{Preliminaries}{Prelims.}
\hi@abbrev{Produce}{Produc.}
\hi@abbrev{Produces}{Producs.}
\hi@abbrev{Production}{Produc.}
\hi@abbrev{Productions}{Producs.}
\hi@abbrev{Reconsideration}{Recons.}
\hi@abbrev{Reconsiderations}{Reconss.}
\hi@abbrev{Record}{R.}
\hi@abbrev{Records}{Rs.}
\hi@abbrev{Rehearing}{Reh'g}
\hi@abbrev{Rehearings}{Reh'gs}
\hi@abbrev{Report and Recommendation}{R. \& R.}
\hi@abbrev{Report and Recommendations}{R. \& Rs.}
\hi@abbrev{Reporter}{Rep.}
\hi@abbrev{Reporters}{Reps.}
\hi@abbrev{Request}{Req.}
\hi@abbrev{Requests}{Reqs.}
\hi@abbrev{Respondent}{Resp't}
\hi@abbrev{Respondents}{Resp'ts}
\hi@abbrev{Response}{Resp.}
\hi@abbrev{Responses}{Resps.}
\hi@abbrev{Summary}{Summ.}
\hi@abbrev{Summaries}{Summs.}
\hi@abbrev{Supplement}{Supp.}
\hi@abbrev{Supplements}{Supps.}
\hi@abbrev{Supplemental}{Supp.}
\hi@abbrev{Supplementals}{Supps.}
\hi@abbrev{Temporary Restraining Order}{TRO}
\hi@abbrev{Temporary Restraining Orders}{TROs}
\hi@abbrev{Testimony}{Test.}
\hi@abbrev{Testimonies}{Tests.}
\hi@abbrev{Transcript}{Tr.}
\hi@abbrev{Transcripts}{Trs.}
\hi@abbrev{Verified Statement}{V.S.}
\hi@abbrev{Verified Statements}{V.Ss.}
\hi@abbrev{Amicus Curiae}{Am. Cur.}
\hi@abbrev{Amici Curiae}{Am. Cur.}
\hi@abbrev{Affidavit}{Aff.}
\hi@abbrev{Affidavits}{Affs.}
\hi@abbrev{Complaint}{Compl.}
\hi@abbrev{Complaints}{Compls.}
\hi@abbrev{Document}{Doc.}
\hi@abbrev{Documents}{Docs.}
\hi@abbrev{Stipulation}{Stip.}
\hi@abbrev{Stipulations}{Stips.}
\hi@abbrev{Stipulated}{Stip.}
\hi@abbrev{Stipulating}{Stip.}
\hi@abbrev{Support}{Supp.}
\hi@abbrev{Supporting}{Supp.}
\hi@abbrev{and}{\&}
\hi@abbrev{The}{}
\hi@abbrev{the}{}
\hi@abbrev{A}{}
\hi@abbrev{a}{}
\hi@abbrev{An}{}
\hi@abbrev{an}{}
\hi@abbrev{of}{}
\hi@abbrev{for}{}
\hi@abbrev{from}{}
\hi@abbrev{in}{}
\hi@abbrev{to}{}
\hi@abbrev{at}{}
\hi@abbrev{by}{}
\hi@abbrev{on}{}
\hi@abbrev{over}{}
\hi@abbrev{per}{}
\hi@abbrev{with}{}
