%
% Document divisions, from Indigo Book Table 13
%
\hi@abbrev{addendum}{add. }{addenda}{adds. }
\hi@abbrev{Addendum}{Add. }{Addenda}{Adds. }
\hi@abbrev{addendum}{add. }{addendums}{adds. }
\hi@abbrev{Addendum}{Add. }{Addendums}{Adds. }
\hi@abbrev{amendment}{amend. }{amendments}{amends. }
\hi@abbrev{Amendment}{Amend. }{Amendments}{Amends. }
\hi@abbrev{annotation}{annot. }{annotations}{annots. }
\hi@abbrev{Annotation}{Annot. }{Annotations}{Annots. }
\hi@abbrev{appendix}{app. }{appendixes}{apps. }
\hi@abbrev{Appendix}{App. }{Appendixes}{Apps. }
\hi@abbrev{appendix}{app. }{appendices}{apps. }
\hi@abbrev{Appendix}{App. }{Appendices}{Apps. }
\hi@abbrev{article}{art. }{articles}{arts. }
\hi@abbrev{Article}{Art. }{Articles}{Arts. }
\hi@abbrev{bibliography}{bibliog. }{bibliographies}{bibliogs. }
\hi@abbrev{Bibliography}{Bibliog. }{Bibliographies}{Bibliogs. }
\hi@abbrev{book}{bk. }{books}{bks. }
\hi@abbrev{Book}{Bk. }{Books}{Bks. }
\hi@abbrev{chapter}{ch. }{chapters}{chs. }
\hi@abbrev{Chapter}{Ch. }{Chapters}{Chs. }
\hi@abbrev{clause}{cl. }{clauses}{cls. }
\hi@abbrev{Clause}{Cl. }{Clauses}{Cls. }
\hi@abbrev{column}{col. }{columns}{cols. }
\hi@abbrev{Column}{Col. }{Columns}{Cols. }
\hi@abbrev{comment}{cmt. }{comments}{cmts. }
\hi@abbrev{Comment}{Cmt. }{Comments}{Cmts. }
\hi@abbrev{commentary}{cmt. }{commentaries}{cmts. }
\hi@abbrev{Commentary}{Cmt. }{Commentaries}{Cmts. }
\hi@abbrev{decision}{dec. }{decisions}{decs. }
\hi@abbrev{Decision}{Dec. }{Decisions}{Decs. }
\hi@abbrev{department}{dept. }{departments}{depts. }
\hi@abbrev{Department}{Dept. }{Departments}{Depts. }
\hi@abbrev{division}{div. }{divisions}{divs. }
\hi@abbrev{Division}{Div. }{Divisions}{Divs. }
\hi@abbrev{example}{ex. }{examples}{exs. }
\hi@abbrev{Example}{Ex. }{Examples}{Exs. }
\hi@abbrev{figure}{fig.}{figures}{figs.}
\hi@abbrev{Figure}{Fig.}{Figures}{Figs.}
\hi@abbrev{folio}{fol. }{folios}{fols. }
\hi@abbrev{Folio}{Fol. }{Folios}{Fols. }
\hi@abbrev{footnote}{n.}{footnotes}{nn.}
\hi@abbrev{Footnote}{N.}{Footnotes}{Nn.}
\hi@abbrev{historical-note}{hist. n.}{historical-notes}{hist. nn.}
\hi@abbrev{Historical-note}{Hist. n.}{Historical-notes}{Hist. nn.}
\hi@abbrev{hypothetical}{hypo. }{hypotheticals}{hypos. }
\hi@abbrev{Hypothetical}{Hypo. }{Hypotheticals}{Hypos. }
\hi@abbrev{illustration}{illus. }{illustrations}{illus. }
\hi@abbrev{Illustration}{Illus. }{Illustrations}{Illus. }
\hi@abbrev{introduction}{intro. }{introductions}{intros. }
\hi@abbrev{Introduction}{Intro. }{Introductions}{Intros. }
\hi@abbrev{line}{l. }{lines}{ll. }
\hi@abbrev{Line}{L. }{Lines}{Ll. }
\hi@abbrev{number}{no. }{numbers}{nos. }
\hi@abbrev{Number}{No. }{Numbers}{Nos. }
\hi@abbrev{part}{pt. }{parts}{pts. }
\hi@abbrev{Part}{Pt. }{Parts}{Pts. }
\hi@abbrev{preamble}{pmbl. }{preambles}{pmbls. }
\hi@abbrev{Preamble}{Pmbl. }{Preambles}{Pmbls. }
\hi@abbrev{principle}{princ. }{principles}{princs. }
\hi@abbrev{Principle}{Princ. }{Principles}{Princs. }
\hi@abbrev{publication}{pub. }{publications}{pubs. }
\hi@abbrev{Publication}{Pub. }{Publications}{Pubs. }
\hi@abbrev{rule}{r. }{rules}{rs. }
\hi@abbrev{Rule}{R. }{Rules}{Rs. }
\hi@abbrev{schedule}{sched. }{schedules}{scheds. }
\hi@abbrev{Schedule}{Sched. }{Schedules}{Scheds. }
\hi@abbrev{serial}{ser. }{serials}{sers. }
\hi@abbrev{Serial}{Ser. }{Serials}{Sers. }
\hi@abbrev{subdivision}{subdiv. }{subdivisions}{subdivs. }
\hi@abbrev{Subdivision}{Subdiv. }{Subdivisions}{Subdivs. }
\hi@abbrev{subsection}{subsec. }{subsections}{subsecs. }
\hi@abbrev{Subsection}{Subsec. }{Subsections}{Subsecs. }
\hi@abbrev{supplement}{supp. }{supplements}{supps. }
\hi@abbrev{Supplement}{Supp. }{Supplements}{Supps. }
\hi@abbrev{table}{tbl.}{tables}{tbls.}
\hi@abbrev{Table}{Tbl.}{Tables}{Tbls.}
\hi@abbrev{title}{tit. }{titles}{tits. }
\hi@abbrev{Title}{Tit. }{Titles}{Tits. }
\hi@abbrev{volume}{vol. }{volumes}{vols. }
\hi@abbrev{Volume}{Vol. }{Volumes}{Vols. }
% Special forms
\hi@abbrev{paragraph}{para. }{paragraphs}{paras. }
\hi@abbrev{section}{sec. }{sections}{secs. }
\hi@abbrev{note}{n.}{notes}{nn.}
% Additional abbreviations
\hi@abbrev{exhibit}{exh. }{exhibits}{exhs. }
\hi@abbrev{Exhibit}{Exh. }{Exhibits}{Exhs. }
\hi@abbrev{claim}{cl. }{claims}{cls. }
\hi@abbrev{Claim}{Cl. }{Claims}{Cls. }
