%
% Abbreviations for legislative documents, from Indigo Book Table 5
%
\hi@abbrev{Annual}{Ann.}
\hi@abbrev{Annuals}{Anns.}
\hi@abbrev{Assembly}{Assemb.}
\hi@abbrev{Assemblies}{Assembs.}
\hi@abbrev{Assemblyman}{Assemb.}
\hi@abbrev{Assemblymans}{Assembs.}
\hi@abbrev{Assemblywoman}{Assemb.}
\hi@abbrev{Assemblywomans}{Assembs.}
\hi@abbrev{Assemblymember}{Assemb.}
\hi@abbrev{Assemblymembers}{Assembs.}
\hi@abbrev{Bill}{B.}
\hi@abbrev{Bills}{Bs.}
\hi@abbrev{Committee}{Comm.}
\hi@abbrev{Committees}{Comms.}
\hi@abbrev{Concurrent}{Con.}
\hi@abbrev{Concurrents}{Cons.}
\hi@abbrev{Conference}{Conf.}
\hi@abbrev{Conferences}{Confs.}
\hi@abbrev{Congress}{Cong.}
\hi@abbrev{Congresses}{Congs.}
\hi@abbrev{Congressional}{Cong.}
\hi@abbrev{Congressionals}{Congs.}
\hi@abbrev{Debate}{Deb.}
\hi@abbrev{Debates}{Debs.}
\hi@abbrev{Delegate}{Del.}
\hi@abbrev{Delegates}{Dels.}
\hi@abbrev{Document}{Doc.}
\hi@abbrev{Documents}{Doc.}
\hi@abbrev{Executive}{Exec.}
\hi@abbrev{Executives}{Execs.}
\hi@abbrev{Federal}{Fed.}
\hi@abbrev{Federals}{Feds.}
\hi@abbrev{House}{H.}
\hi@abbrev{Houses}{Hs.}
\hi@abbrev{House of Delegates}{H.D.}
\hi@abbrev{House of Representatives}{H.R.}
\hi@abbrev{Joint}{J.}
\hi@abbrev{Joints}{Js.}
\hi@abbrev{Legislation}{Legis.}
\hi@abbrev{Legislations}{Legiss.}
\hi@abbrev{Legislative}{Legis.}
\hi@abbrev{Legislatives}{Legiss.}
\hi@abbrev{Legislature}{Leg.}
\hi@abbrev{Legislatures}{Legs.}
\hi@abbrev{Miscellaneous}{Misc.}
\hi@abbrev{Miscellaneouses}{Miscs.}
\hi@abbrev{Number}{No.}
\hi@abbrev{Numbers}{Nos.}
\hi@abbrev{Record}{Rec.}
\hi@abbrev{Records}{Recs.}
\hi@abbrev{Register}{Reg.}
\hi@abbrev{Registers}{Regs.}
\hi@abbrev{Regular}{Reg.}
\hi@abbrev{Regulars}{Regs.}
\hi@abbrev{Report}{Rep.}
\hi@abbrev{Reports}{Reps.}
\hi@abbrev{Representative}{Rep.}
\hi@abbrev{Representatives}{Reps.}
\hi@abbrev{Resolution}{Res.}
\hi@abbrev{Resolutions}{Ress.}
\hi@abbrev{Senate}{S.}
\hi@abbrev{Senates}{Ss.}
\hi@abbrev{Senator}{Sen.}
\hi@abbrev{Senators}{Sens.}
\hi@abbrev{Service}{Serv.}
\hi@abbrev{Services}{Servs.}
\hi@abbrev{Session}{Sess.}
\hi@abbrev{Sessions}{Sesss.}
\hi@abbrev{Special}{Spec.}
\hi@abbrev{Specials}{Specs.}
\hi@abbrev{Subcommittee}{Subcomm.}
\hi@abbrev{Subcommittees}{Subcomms.}
\hi@abbrev{and}{\&}
\hi@abbrev{The}{}
\hi@abbrev{the}{}
\hi@abbrev{A}{}
\hi@abbrev{a}{}
\hi@abbrev{An}{}
\hi@abbrev{an}{}
\hi@abbrev{of}{}
\hi@abbrev{for}{}
\hi@abbrev{from}{}
\hi@abbrev{in}{}
\hi@abbrev{to}{}
\hi@abbrev{at}{}
\hi@abbrev{by}{}
\hi@abbrev{on}{}
\hi@abbrev{over}{}
\hi@abbrev{per}{}
\hi@abbrev{with}{}
