\begin{enumerate}
\item Abbie Mills lends her friend Ichabod Crane a book when Ichabod is visiting
her home.  A week later, she calls Ichabod and tells him to keep the book. 
Does he own the book?  At what point did he come to own the book?

\item Walter Bishop, on his deathbed, tells Olivia Dunham, ``Because I love you
like a daughter and I want to ensure your financial security, I want you to
have my house in Disturbia City.  Continue to rent it out as I have done and
you will always have an income.''  He hands her the deed to the house, then
dies.  Walter's will leaves all his property to Peter Bishop.  For the last
five years, the property has been rented to Astrid Farnsworth as a furnished
dwelling.  Peter Bishop sues Dunham for possession of the furniture in the
house as well as the rare book collection stored on the bookshelves of the
house.  Who owns the furniture, and why?  Who owns the rare book collection,
and why?


\item Victoria Grayson hires Amanda Clarke as nanny to her children, Charlotte
and Daniel Grayson.  One day, Clarke praises one of Victoria Grayson's diamond
necklaces.  Grayson says, ``You can have it if Charlotte and Daniel are
accepted into Evernight Academy.  In the meantime, you can wear it.''  Clarke
thanks Grayson and wears the necklace thereafter.  Right before the Evernight
entrance results are released, Grayson fires Clarke, telling her, ``I despise
you and I will never do anything for you!''  Clarke leaves, taking the necklace
with her.  A day later, Charlotte and Daniel are accepted into Evernight.  A
day after that, Victoria Grayson sends a letter to Clarke demanding the return
of the necklace.  Who owns the necklace and why?


\item Damon and Stefan Salvatore, brothers, owned their family home.  They
fought, and, in an attempt to placate Stefan, Damon executed a deed
transferring his ownership interest in the home to Stefan.  Damon went to
Stefan's study and handed him the deed.  Stefan dropped the deed (it landed on
his desk) and yelled, ``You'll never buy my forgiveness!'' Then he punched
Damon and left town.  Damon continued to live in the house.  Two years later,
with Stefan still gone, Damon executed a new deed transferring his ownership
interest to Elena Gilbert, who had no idea about the previous deed.  Damon
handed the deed to Elena, telling her, ``With all my family has done to yours,
this is the least I can do.''  She thanked him.  A year later, Stefan returned.
 Damon's first deed was still on Stefan's desk, albeit very dusty.  At that
point, Stefan put the deed in his safety deposit box at the bank and thanked
Damon for the gesture.  Damon said, ``No problem.''  Which gift was completed
first?


\item Fox Mulder gives Dana Scully a keychain with an Apollo 11 medallion on it.
 But when she tries to put her keys on it, the ring doesn't work.  He tells
her, ``Let me have that fixed for you,'' and she lets him have the keychain. 
Afterwards, he decides that he doesn't want her to have the keychain.  Who owns
the keychain?
\end{enumerate}

