\reading{The Fair Housing Act of 1968}

\readingcite{42 U.S.C. \S\S~3601--3619}

\S~3601.  \textsc{Declaration of Policy}.---It is the policy of the United
States to provide, within constitutional limitations, for fair housing
throughout the United States.

\S~3602.  \textsc{Definitions}.---As used in this subchapter\ldots
\begin{statute}
\item (c) ``Family'' includes a single individual.\ldots

\item (h) ``Handicap'' means, with respect to a person---

\begin{statute}
\item (1) a physical or mental impairment which substantially limits one or
more of such person's major life activities,

\item (2) a record of having such an impairment, or

\item (3) being regarded as having such an impairment,
\end{statute}
but such term does not
include current, illegal use of or addiction to a controlled
substance\edfootnote{``Controlled substance'' is defined generally to
include regulated drugs but not alcohol.}\ldots.
%(as defined in section 802 of title 21).\ldots

\item (k) ``Familial status'' means one or more individuals (who
have not attained the age of 18 years) being domiciled with---
\begin{statute}
\item (1) a parent or another person having legal custody of such individual or
individuals; or  

\item (2) the designee of such parent or other person having such custody, with
the written permission of such parent or other person. 
\end{statute}
The protections afforded against discrimination on the basis of familial status
shall apply to any person who is pregnant or is in the process of securing
legal custody of any individual who has not attained the age of 18 years.
\end{statute}


\S~3604. \textsc{Discrimination in the sale or rental of housing and other
prohibited practices.}---[Subject to certain exceptions], it shall be
unlawful---


\begin{statute}
\item (a) To refuse to sell or rent after the making of a bona fide offer, or to
refuse to negotiate for the sale or rental of, or otherwise make unavailable or
deny, a dwelling to any person because of race, color, religion, sex, familial
status, or national origin [hereinafter RCRSFSNO].

\item (b) To discriminate against any person in the terms, conditions, or
privileges of sale or rental of a dwelling, or in the provision of services or
facilities in connection therewith, because of [RCRSFSNO].

\item (c) To [publish an]
%make, print, or publish, or cause to be made, printed, or published
%any notice, statement, or
advertisement, with respect to the sale or rental of a
dwelling that indicates any preference, limitation, or discrimination based on
[RCRSFSNO or handicap], or an intention to make any such preference, limitation,
or discrimination.

\item (d) To represent to any person because of [RCRSFSNO or handicap] that any
dwelling is not available for inspection, sale, or rental when such dwelling is
in fact so available.

\item (e) For profit, to induce or attempt to induce any person to sell or rent
any dwelling by representations regarding the entry or prospective entry into
the neighborhood of a person or persons of a particular [RCRSFSNO or handicap].

\end{statute}

\S~3603(b). Nothing in section 3604 of this title (other than subsection (c))
shall apply to---
\begin{statute}
\item (1) any single-family house sold or rented by an owner [so long as the
owner meets a variety of tests intended to distinguish a ``private individual
owner'' from a commercial real estate dealer]; or
\item (2) rooms or units in dwellings containing living quarters occupied or
intended to be occupied by no more than four families living independently of
each other, if the owner actually maintains and occupies one of such living
quarters as his residence.
\end{statute}
