\expected{jacque-v-steenberg}

\item \label{bkm:Ref328040570}Would (or should) the result in \textit{Jacque}
have been different if, instead of a mobile home seller making a scheduled
delivery to a customer, the defendant had been an ambulance company responding
to a call of a suspected heart attack? Of a broken leg? What if the
snow-covered private road had instead been a recently collapsed bridge? What if
Steenberg had tried to take the road despite the risks, and the truck had
accidentally tipped and fallen onto the Jacques' land?

\item \label{bkm:Ref328040574}Would (or should) the result in \textit{Jacque}
have been different if, instead of steadfastly refusing to permit Steenberg's
delivery truck to cross their land, the Jacques had demanded a large sum of
money as a condition of permitting the crossing, which Steenberg refused to
pay? Would the ultimate monetary award have been different? If so, what
incentive does this case give property owners facing requests from third
parties for the use of their otherwise idle resources? Would Steenberg have
been better off not asking permission in the first place?

\item Blackstone's description of ``that sole and despotic dominion which one
man claims and exercises over the external things of the world, in total
exclusion of the right of any other individual in the universe'' is one of
the most famous---and quotable---definitions of property ever written in
English. But it is also widely acknowledged to be hyperbolic to the point of
falsity. Can you see why? What aspects of Blackstone's own discussion of the
``absolute right'' of property are inconsistent with the ``total exclusion of
the right of any other individual in the universe''?

\item Would we really want our system of property to give private owners such
``sole and despotic dominion\ldots over the external things of the world''? The
kind of dominion exercised by the Jacques? No matter what? Consider this: what
kinds of problems could a motivated and unscrupulous property owner armed with
such awesome power cause?


